---
title: "BHAMM Sex Differences Segmented Regression Study"
subtitle: "Script 2 - Analyses Spatial Context Memory Task"
author: "Annalise LaPlume"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    highlight: textmate
    theme: paper
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

# Data setup

## Load data file and packages
```{r setup, include=TRUE, warning=FALSE, message=FALSE}

# Install and load packages
if(!require("haven"))install.packages("haven") 
if(!require("psych"))install.packages("psych") 
if(!require("dplyr"))install.packages("dplyr") 
if(!require("segmented"))install.packages("segmented") 
if(!require("ggplot2"))install.packages("ggplot2") 
if(!require("janitor"))install.packages("janitor") 
if(!require("forcats"))install.packages("forcats") 

library(haven) # load spss files
library(psych) # basic statistics functions
library(dplyr) # basic statistics functions
library(segmented) # segmented regression
library(ggplot2) # plot results
library(janitor) # examining and cleaning data
library(forcats) # working with categorical variables


# Set working directory to current folder
# Edit to specify where datafile is located
setwd("C:/Users/annal/OneDrive/Documents/BHAMM/Segmented Regression/Sex Differences")

# Import data
bhamm <- read_sav("BHAMM_Behav_Data_For_Analysis_2024_03.sav")

# view data sample
head(bhamm)

names(bhamm)
```

## Dependent variables

```{r dvs, include=TRUE, warning=FALSE, message=FALSE}
# DV
# cs_rate_easy: Easy Correct Source Accuracy (OLD LOCATION = OLD LOCATION)
# cs_rate_hard: Hard Correct Source Accuracy (OLD LOCATION = OLD LOCATION)
# recog_rate_easy: Easy Rate Recognition (OLD = FAMILIAR, NO LOCATION)
# recog_rate_hard:Hard Rate Recognition (OLD = FAMILIAR, NO LOCATION)

## Check variable structure
str(bhamm$cs_rate_easy)
str(bhamm$cs_rate_hard)
str(bhamm$recog_rate_easy)
str(bhamm$recog_rate_hard)

str(bhamm$rt_cs_mean)
str(bhamm$rt_cs_hard)
str(bhamm$rt_recog_mean)
str(bhamm$rt_recog_hard)

# Examine variable values
describe(bhamm$cs_rate_easy)
describe(bhamm$cs_rate_hard)
describe(bhamm$recog_rate_easy)
describe(bhamm$recog_rate_hard)

# Create percentages
bhamm$cs_easy_percent <- bhamm$cs_rate_easy*100
bhamm$cs_hard_percent <- bhamm$cs_rate_hard*100
bhamm$recog_easy_percent <- bhamm$recog_rate_easy*100
bhamm$recog_hard_percent <- bhamm$recog_rate_hard*100

# calculate accuracy average for easy and hard conditions
bhamm$cs_mean <- rowMeans(bhamm[,c('cs_easy_percent', 
                                   'cs_hard_percent')],
                          na.rm=TRUE)

bhamm$recog_mean <- rowMeans(bhamm[,c('recog_easy_percent', 
                                      'recog_hard_percent')],
                             na.rm=TRUE)


# calculate rt average for easy and hard conditions
bhamm$rt_cs_mean <- rowMeans(bhamm[,c('rt_cs_easy', 
                                   'rt_cs_hard')],
                          na.rm=TRUE)

bhamm$rt_recog_mean <- rowMeans(bhamm[,c('rt_recog_easy', 
                                      'rt_recog_hard')],
                             na.rm=TRUE)

describe(bhamm$cs_mean)
describe(bhamm$recog_mean)
describe(bhamm$rt_cs_mean)
describe(bhamm$rt_recog_mean)


```

```{r boundaries, include=TRUE, warning=FALSE, message=FALSE}

# Examine women in overlapping ages of meno groups

bhamm$s2_meno_group <- as.factor(bhamm$s2_meno_group)

# Count women in pre and post menopause
table(bhamm$s2_meno_group)

# Count premenopause (meno group=1) age >50
dplyr::count(subset(bhamm, s2_meno_group == "1" & s2_age >50)) # n=1

# Count postmenopause (meno group=3) age <50
dplyr::count(subset(bhamm,s2_meno_group == "3" & s2_age<50)) # n=1

# One premeno woman age 53, and one postmeno woman age 47
# Aside from these two women, no other overlap in ages between pre and post meno
```

## Independent variables
```{r ivs, include=TRUE, warning=FALSE, message=FALSE}
# Convert sex and menopause status into factors
bhamm$sex <- as.factor(bhamm$sex)
bhamm$s2_meno_group <- as.factor(bhamm$s2_meno_group)

levels(bhamm$s2_meno_group)
levels(bhamm$s2_meno_group) <- c("Premenopause", "Postmenopause")
levels(bhamm$s2_meno_group)

# Count n per menopause group
table(bhamm$s2_meno_group)

# Sex
str(bhamm$sex)
bhamm$sex <- as.factor(bhamm$sex)
levels(bhamm$sex) <- c("Male", "Female")
table(bhamm$sex)

# Add NA as a level to meno group factor 
bhamm$s2_meno_group_na <- addNA(bhamm$s2_meno_group)

# Sex by menopause status
table(bhamm$sex, bhamm$s2_meno_group_na)

# Age
str(bhamm$s2_age)
describe(bhamm$s2_age)

```

## Sex differences factor
```{r sex diffs f, include=TRUE, warning=FALSE, message=FALSE}

# Check structure of variables
str(bhamm$s2_meno_group_na)
str(bhamm$sex)

# Convert sex to factor
bhamm$sex_f <- as.factor(bhamm$sex)
str(bhamm$sex_f)

# Convert sex and menopause status into a single factor
bhamm$sex_meno_group <- fct_cross(bhamm$sex_f, 
                                   bhamm$s2_meno_group_na, 
                                   keep_empty = TRUE)

str(bhamm$sex_meno_group)
levels(bhamm$sex_meno_group)

# Collapse into single factor for males
bhamm$sex_meno <- fct_collapse(bhamm$sex_meno_group, 
                                      Male = c("Male:Premenopause",
                                                "Male:Postmenopause",
                                                "Male:NA"))
# Exclude NA females
bhamm <- subset(bhamm, sex_meno != "Female:NA")

table(bhamm$sex_meno)
```

## Separate datasets for males and females

```{r males females, include=TRUE, warning=FALSE, message=FALSE}
bhamm_m <- subset(bhamm, sex == "Male")
dplyr::count(bhamm_m)

bhamm_f <- subset(bhamm, sex == "Female")
dplyr::count(bhamm_f)
```




# Scattterplots
Plot scatterplot of dependent variables by age 
with separate lines for sex & menopause group

## Accuracy 
```{r acc scatterplots, include=TRUE, warning=FALSE, message=FALSE}
# Correct source easy 

# save plot in jpeg format in current directory
jpeg(file="cs_scatterplot_bhamm.jpeg", width = 4, height = 4, 
     units = 'in', res = 300)

ggplot(data=bhamm,                                         
       aes(x=s2_age, y = cs_mean,                                       
           colour=sex_meno,                                        
           na.rm= TRUE)) +                                              
  ylim (0,100) +
  geom_point(size = 2, alpha = .6,                                     
             position = position_jitter(height = 0.1)) +  
  labs(x="Age", y="Accuracy (% correct)",
       colour="",                   
       title="Spatial Context Memory Task \n Correct Spatial Context Retrieval") +
      theme_classic() +                                            
    theme(axis.text=element_text(size=14),                          
          axis.title=element_text(size=14),                        
           plot.title=element_text(size=14, face="bold", hjust = 0.5),
          legend.text=element_text(size=10),
          legend.position="none") +
           geom_smooth(method = "lm") +
  scale_colour_manual(values=c("cornflowerblue", "brown1", "darkorange1"))

# call this function to save the file 
dev.off()

# Recognition easy by age
jpeg(file="recog_scatterplot_bhamm.jpeg", width = 4, height = 4, 
     units = 'in', res = 300)

ggplot(data=bhamm,                                         
       aes(x=s2_age, y = recog_mean,                                    
           colour=sex_meno,                                        
           na.rm= TRUE)) +                                              
  ylim (0,100) +
  geom_point(size = 2, alpha = .6,                                     
             position = position_jitter(height = 0.1)) +  
  labs(x="Age", y="Accuracy (% correct)",
       colour="",                   
       title="Spatial Context Memory Task \n Correct Recognition") +              
      theme_classic() +                                            
    theme(axis.text=element_text(size=14),                          
          axis.title=element_text(size=14),                        
           plot.title=element_text(size=14, face="bold", hjust = 0.5),
          legend.text=element_text(size=10),
          legend.position="none") +
           geom_smooth(method = "lm")+
  scale_colour_manual(values=c("cornflowerblue", "brown1", "darkorange1"))

dev.off()

# Image with legend (crop and use for legend)

jpeg(file="legend.jpeg", width = 4, height = 4, 
     units = 'in', res = 300)

ggplot(data=bhamm,                                         
       aes(x=s2_age, y = cs_mean,                                       
           colour=sex_meno,                                        
           na.rm= TRUE)) +                                              
  ylim (0,100) +
  geom_point(size = 2, alpha = .6,                                     
             position = position_jitter(height = 0.1)) +  
  labs(x="", y="",
       colour="",                   
       title="") +                
      theme_classic() +                                            
    theme(axis.text=element_text(size=1),                          
          axis.title=element_text(size=1),                        
           plot.title=element_text(size=2, face="bold",            
                                    hjust = 0.5),
          legend.text=element_text(size=16),
          legend.position="right") +
           geom_smooth(method = "lm")+
  scale_colour_manual(values=c("cornflowerblue", "brown1", "darkorange1"))

dev.off()

```


## Reaction times 

```{r rt scatterplots, include=TRUE, warning=FALSE, message=FALSE}
# Correct source easy 

# save plot in jpeg format in current directory
jpeg(file="cs_scatterplot_rt_bhamm.jpeg", width = 4, height = 4, 
     units = 'in', res = 300)

ggplot(data=bhamm,                                         
       aes(x=s2_age, y = rt_cs_mean,                                       
           colour=sex_meno,                                        
           na.rm= TRUE)) +                                              
  geom_point(size = 2, alpha = .6,                                     
             position = position_jitter(height = 0.1)) +  
  ylim(0,4000) +
  labs(x="Age", y="Reaction time (ms)",
       colour="",                   
       title="Spatial Context Memory Task \n Correct Spatial Context Retrieval") +
      theme_classic() +                                            
    theme(axis.text=element_text(size=14),                          
          axis.title=element_text(size=14),                        
           plot.title=element_text(size=14, face="bold", hjust = 0.5),
          legend.text=element_text(size=10),
          legend.position="none") +
           geom_smooth(method = "lm")+
  scale_colour_manual(values=c("cornflowerblue", "brown1", "darkorange1"))

# call this function to save the file 
dev.off()


# Recognition easy by age
jpeg(file="recog_scatterplot_rt_bhamm.jpeg", width = 4, height = 4, 
     units = 'in', res = 300)

ggplot(data=bhamm,                                         
       aes(x=s2_age, y = rt_recog_mean,                                       
           colour=sex_meno,                                        
           na.rm= TRUE)) +                                              
  geom_point(size = 2,  alpha = .6,                                     
             position = position_jitter(height = 0.1)) +  
  ylim(0,6500) +
  labs(x="Age", y="Reaction time (ms)",
       colour="",                   
       title="Spatial Context Memory Task \n Correct Recognition") +              
      theme_classic() +                                            
    theme(axis.text=element_text(size=14),                          
          axis.title=element_text(size=14),                        
           plot.title=element_text(size=14, face="bold", hjust = 0.5),
          legend.text=element_text(size=10),
          legend.position="none") +
           geom_smooth(method = "lm")+
  scale_colour_manual(values=c("cornflowerblue", "brown1", "darkorange1"))

dev.off()
```


# Segmented regression 

Separate segmented models for males and females

## Accuracy

### Correct source - Males

```{r segmented cs acc m, include=TRUE, warning=FALSE, message=FALSE}

# linear model
cs_mean_m_lm <-lm(cs_mean ~ s2_age + edu, data=bhamm_m)
summary(cs_mean_m_lm)

# test for breakpoint
davies.test(cs_mean_m_lm, ~ s2_age)
pscore.test(cs_mean_m_lm, ~ s2_age)

# segmented model with one breakpoint
cs_mean_m_seg_1 <- segmented(cs_mean_m_lm, seg.Z=~s2_age)

summary(cs_mean_m_seg_1) 

# segmented model with two breakpoints
cs_mean_m_seg_2 <- segmented(cs_mean_m_lm, seg.Z=~s2_age, 
                           npsi=2, control=seg.control(display=FALSE))

summary.segmented(cs_mean_m_seg_2)


# Model comparison: lm vs 1 segment vs 2 segments
anova(cs_mean_m_lm, cs_mean_m_seg_1, cs_mean_m_seg_2)
AIC(cs_mean_m_lm, cs_mean_m_seg_1, cs_mean_m_seg_2)
BIC(cs_mean_m_lm, cs_mean_m_seg_1, cs_mean_m_seg_2)
```

### Correct source - Females

```{r segmented cs acc f, include=TRUE, warning=FALSE, message=FALSE}

# linear model
cs_mean_f_lm <-lm(cs_mean ~ s2_age + edu, data=bhamm_f)
summary(cs_mean_f_lm)

# test for breakpoint
davies.test(cs_mean_f_lm, ~ s2_age)
pscore.test(cs_mean_f_lm, ~ s2_age)

# segmented model with one breakpoint
cs_mean_f_seg_1 <- segmented(cs_mean_f_lm, seg.Z=~s2_age)

summary(cs_mean_f_seg_1) 

# segmented model with two breakpoints
cs_mean_f_seg_2 <- segmented(cs_mean_f_lm, seg.Z=~s2_age, 
                           npsi=2, control=seg.control(display=FALSE))

summary.segmented(cs_mean_f_seg_2)

# Model comparison: lm vs 1 segment vs 2 segments
anova(cs_mean_f_lm, cs_mean_f_seg_1, cs_mean_f_seg_2)
AIC(cs_mean_f_lm, cs_mean_f_seg_1, cs_mean_f_seg_2)
BIC(cs_mean_f_lm, cs_mean_f_seg_1, cs_mean_f_seg_2)

# slope
slope(cs_mean_f_seg_1)

# confidence intervals
confint(cs_mean_f_seg_1)

# plot
# plot.segmented(cs_mean_f_seg_1)
```

### Recognition - Males

```{r segmented recog acc m, include=TRUE, warning=FALSE, message=FALSE}
# linear model
recog_mean_m_lm <-lm(recog_mean ~ s2_age + edu, data=bhamm_m)
summary(recog_mean_m_lm)

# test for breakpoint
davies.test(recog_mean_m_lm, ~ s2_age)
pscore.test(recog_mean_m_lm, ~ s2_age)

# segmented model with one breakpoint
recog_mean_m_seg_1 <- segmented(recog_mean_m_lm, seg.Z=~s2_age)
summary(recog_mean_m_seg_1) 

# segmented model with two breakpoints
recog_mean_m_seg_2 <- segmented(recog_mean_m_lm, seg.Z=~s2_age, 
                           npsi=2, control=seg.control(display=FALSE))
summary.segmented(recog_mean_m_seg_2)

# Model comparison: lm vs 1 segment vs 2 segments
anova(recog_mean_m_lm, recog_mean_m_seg_1, recog_mean_m_seg_2)
AIC(recog_mean_m_lm, recog_mean_m_seg_1, recog_mean_m_seg_2)
BIC(recog_mean_m_lm, recog_mean_m_seg_1, recog_mean_m_seg_2)


```

### Recognition - Females

```{r segmented recog acc f, include=TRUE, warning=FALSE, message=FALSE}
# linear model
recog_mean_f_lm <-lm(recog_mean ~ s2_age + edu, data=bhamm_f)
summary(recog_mean_f_lm)

# test for breakpoint
davies.test(recog_mean_f_lm, ~ s2_age)
pscore.test(recog_mean_f_lm, ~ s2_age)

# segmented model with one breakpoint
recog_mean_f_seg_1 <- segmented(recog_mean_f_lm, seg.Z=~s2_age)
summary(recog_mean_f_seg_1) 

# segmented model with two breakpoints
recog_mean_f_seg_2 <- segmented(recog_mean_f_lm, seg.Z=~s2_age, 
                           npsi=2, control=seg.control(display=FALSE))
summary.segmented(recog_mean_f_seg_2)

# Model comparison: lm vs 1 segment vs 2 segments
anova(recog_mean_f_lm, recog_mean_f_seg_1, recog_mean_f_seg_2)
AIC(recog_mean_f_lm, recog_mean_f_seg_1, recog_mean_f_seg_2)
BIC(recog_mean_f_lm, recog_mean_f_seg_1, recog_mean_f_seg_2)

# slope
slope(recog_mean_f_seg_1)

# confidence intervals
confint(recog_mean_f_seg_1)

# plot
# plot.segmented(recog_mean_f_seg_1)
```

## Reaction times

### Correct source - Males

```{r segmented rt cs m, include=TRUE, warning=FALSE, message=FALSE}

# linear model
rt_cs_mean_m_lm <-lm(rt_cs_mean ~ s2_age + edu, data=bhamm_m)
summary(rt_cs_mean_m_lm)

# test for breakpoint
davies.test(rt_cs_mean_m_lm, ~ s2_age)
pscore.test(rt_cs_mean_m_lm, ~ s2_age)

# segmented model with one breakpoint
rt_cs_mean_m_seg_1 <- segmented(rt_cs_mean_m_lm, seg.Z=~s2_age)

summary(rt_cs_mean_m_seg_1) 

# segmented model with two breakpoints
rt_cs_mean_m_seg_2 <- segmented(rt_cs_mean_m_lm, seg.Z=~s2_age, 
                           npsi=2, control=seg.control(display=FALSE))

summary.segmented(rt_cs_mean_m_seg_2)


# Model comparison: lm vs 1 segment vs 2 segments
anova(rt_cs_mean_m_lm, rt_cs_mean_m_seg_1, rt_cs_mean_m_seg_2)
AIC(rt_cs_mean_m_lm, rt_cs_mean_m_seg_1, rt_cs_mean_m_seg_2)
BIC(rt_cs_mean_m_lm, rt_cs_mean_m_seg_1, rt_cs_mean_m_seg_2)
```

### Correct source - Females

```{r segmented rt_cs f, include=TRUE, warning=FALSE, message=FALSE}

# linear model
rt_cs_mean_f_lm <-lm(rt_cs_mean ~ s2_age + edu, data=bhamm_f)
summary(rt_cs_mean_f_lm)

# test for breakpoint
davies.test(rt_cs_mean_f_lm, ~ s2_age)
pscore.test(rt_cs_mean_f_lm, ~ s2_age)

# segmented model with one breakpoint
rt_cs_mean_f_seg_1 <- segmented(rt_cs_mean_f_lm, seg.Z=~s2_age)

summary(rt_cs_mean_f_seg_1) 

# segmented model with two breakpoints
rt_cs_mean_f_seg_2 <- segmented(rt_cs_mean_f_lm, seg.Z=~s2_age, 
                           npsi=2, control=seg.control(display=FALSE))

summary.segmented(rt_cs_mean_f_seg_2)

# Model comparison: lm vs 1 segment vs 2 segments
anova(rt_cs_mean_f_lm, rt_cs_mean_f_seg_1, rt_cs_mean_f_seg_2)
AIC(rt_cs_mean_f_lm, rt_cs_mean_f_seg_1, rt_cs_mean_f_seg_2)
BIC(rt_cs_mean_f_lm, rt_cs_mean_f_seg_1, rt_cs_mean_f_seg_2)
```

### Recognition - Males

```{r segmented rt recog m, include=TRUE, warning=FALSE, message=FALSE}
# linear model
rt_recog_mean_m_lm <-lm(rt_recog_mean ~ s2_age + edu, data=bhamm_m)
summary(rt_recog_mean_m_lm)

# test for breakpoint
davies.test(rt_recog_mean_m_lm, ~ s2_age)
pscore.test(rt_recog_mean_m_lm, ~ s2_age)

# segmented model with one breakpoint
rt_recog_mean_m_seg_1 <- segmented(rt_recog_mean_m_lm, seg.Z=~s2_age)
summary(rt_recog_mean_m_seg_1) 

# segmented model with two breakpoints
rt_recog_mean_m_seg_2 <- segmented(rt_recog_mean_m_lm, seg.Z=~s2_age, 
                           npsi=2, control=seg.control(display=FALSE))
summary.segmented(rt_recog_mean_m_seg_2)

# Model comparison: lm vs 1 segment vs 2 segments
anova(rt_recog_mean_m_lm, rt_recog_mean_m_seg_1, rt_recog_mean_m_seg_2)
AIC(rt_recog_mean_m_lm, rt_recog_mean_m_seg_1, rt_recog_mean_m_seg_2)
BIC(rt_recog_mean_m_lm, rt_recog_mean_m_seg_1, rt_recog_mean_m_seg_2)
```

### Recognition - Females

```{r segmented rt recog f, include=TRUE, warning=FALSE, message=FALSE}
# linear model
rt_recog_mean_f_lm <-lm(rt_recog_mean ~ s2_age + edu, data=bhamm_f)
summary(rt_recog_mean_f_lm)

# test for breakpoint
davies.test(rt_recog_mean_f_lm, ~ s2_age)
pscore.test(rt_recog_mean_f_lm, ~ s2_age)

# segmented model with one breakpoint
rt_recog_mean_f_seg_1 <- segmented(rt_recog_mean_f_lm, seg.Z=~s2_age)
summary(rt_recog_mean_f_seg_1) 

# segmented model with two breakpoints
rt_recog_mean_f_seg_2 <- segmented(rt_recog_mean_f_lm, seg.Z=~s2_age, 
                           npsi=2, control=seg.control(display=FALSE))
summary.segmented(rt_recog_mean_f_seg_2)

# Model comparison: lm vs 1 segment vs 2 segments
anova(rt_recog_mean_f_lm, rt_recog_mean_f_seg_1, rt_recog_mean_f_seg_2)
AIC(rt_recog_mean_f_lm, rt_recog_mean_f_seg_1, rt_recog_mean_f_seg_2)
BIC(rt_recog_mean_f_lm, rt_recog_mean_f_seg_1, rt_recog_mean_f_seg_2)
```


# Summary of results 

## Accuracy

Hierarchical model comparison
```{r segmented accuracy model comparison, include=TRUE, warning=FALSE, message=FALSE}
# Table 3

# CS

# p-values (chi-square test)
anova(cs_mean_m_lm, cs_mean_m_seg_1, cs_mean_m_seg_2) 
anova(cs_mean_f_lm, cs_mean_f_seg_1, cs_mean_f_seg_2)

# change in AIC
AIC(cs_mean_m_seg_1) - AIC (cs_mean_m_lm)
AIC(cs_mean_m_seg_2) - AIC (cs_mean_m_seg_1)

AIC(cs_mean_f_seg_1) - AIC (cs_mean_f_lm)
AIC(cs_mean_f_seg_2) - AIC (cs_mean_f_seg_1)

# change in r^2
summary(cs_mean_m_seg_1)$adj.r.squared - summary(cs_mean_m_lm)$adj.r.squared
summary(cs_mean_m_seg_2)$adj.r.squared - summary(cs_mean_m_seg_1)$adj.r.squared

summary(cs_mean_f_seg_1)$adj.r.squared - summary(cs_mean_f_lm)$adj.r.squared
summary(cs_mean_f_seg_2)$adj.r.squared - summary(cs_mean_f_seg_1)$adj.r.squared

# Recognition
anova(recog_mean_m_lm, recog_mean_m_seg_1, recog_mean_m_seg_2)
anova(recog_mean_f_lm, recog_mean_f_seg_1, recog_mean_f_seg_2)

AIC(recog_mean_m_seg_1) - AIC (recog_mean_m_lm)
AIC(recog_mean_m_seg_2) - AIC (recog_mean_m_seg_1)

AIC(recog_mean_f_seg_1) - AIC (recog_mean_f_lm)
AIC(recog_mean_f_seg_2) - AIC (recog_mean_f_seg_1)

summary(recog_mean_m_seg_1)$adj.r.squared - summary(recog_mean_m_lm)$adj.r.squared
summary(recog_mean_m_seg_2)$adj.r.squared - summary(recog_mean_m_seg_1)$adj.r.squared

summary(recog_mean_f_seg_1)$adj.r.squared - summary(recog_mean_f_lm)$adj.r.squared
summary(recog_mean_f_seg_2)$adj.r.squared - summary(recog_mean_f_seg_1)$adj.r.squared
```

Estimates from best fitting models

```{r accuracy fitted estimates, include=TRUE, warning=FALSE, message=FALSE}
# Slopes
summary(cs_mean_m_lm)

slope(cs_mean_f_seg_1) 
# B1: t(52) = -0.54, p = .59
# B2: t(52) = -2.59, p = .01

summary(recog_mean_m_lm)

slope(recog_mean_f_seg_1)
# B1: t(52) = 0.76, p = .45
# B2: t(52) = 2.08, p = .04


# For segmented models, get p-values for slopes from t-statistics using online calculator
# https://www.socscistatistics.com/pvalues/tdistribution.aspx
# 2-tailed test and 52 df

# Changepoints
confint(cs_mean_f_seg_1)

confint(recog_mean_f_seg_1)
```

## Reaction times
```{r segmented final rt, include=TRUE, warning=FALSE, message=FALSE}

# CS
# p-values (chi-square test)
anova(rt_cs_mean_m_lm, rt_cs_mean_m_seg_1, rt_cs_mean_m_seg_2)
anova(rt_cs_mean_f_lm, rt_cs_mean_f_seg_1, rt_cs_mean_f_seg_2)

# change in AIC
AIC(rt_cs_mean_m_seg_1) - AIC (rt_cs_mean_m_lm)
AIC(rt_cs_mean_m_seg_2) - AIC (rt_cs_mean_m_seg_1)

AIC(rt_cs_mean_f_seg_1) - AIC (rt_cs_mean_f_lm)
AIC(rt_cs_mean_f_seg_2) - AIC (rt_cs_mean_f_seg_1)

# change in r^2
summary(rt_cs_mean_m_seg_1)$adj.r.squared - summary(rt_cs_mean_m_lm)$adj.r.squared
summary(rt_cs_mean_m_seg_2)$adj.r.squared - summary(rt_cs_mean_m_seg_1)$adj.r.squared

summary(rt_cs_mean_f_seg_1)$adj.r.squared - summary(rt_cs_mean_f_lm)$adj.r.squared
summary(rt_cs_mean_f_seg_2)$adj.r.squared - summary(rt_cs_mean_f_seg_1)$adj.r.squared


# Recognition
anova(rt_recog_mean_m_lm, rt_recog_mean_m_seg_1, rt_recog_mean_m_seg_2)
anova(rt_recog_mean_f_lm, rt_recog_mean_f_seg_1, rt_recog_mean_f_seg_2)

AIC(rt_recog_mean_m_seg_1) - AIC(rt_recog_mean_m_lm)
AIC(rt_recog_mean_m_seg_2) - AIC(rt_recog_mean_m_seg_1)

AIC(rt_recog_mean_f_seg_1) - AIC(rt_recog_mean_f_lm)
AIC(rt_recog_mean_f_seg_2) - AIC(rt_recog_mean_f_seg_1)

summary(rt_recog_mean_m_seg_1)$adj.r.squared - summary(rt_recog_mean_m_lm)$adj.r.squared
summary(rt_recog_mean_m_seg_2)$adj.r.squared - summary(rt_recog_mean_m_seg_1)$adj.r.squared

summary(rt_recog_mean_f_seg_1)$adj.r.squared - summary(rt_recog_mean_f_lm)$adj.r.squared
summary(rt_recog_mean_f_seg_2)$adj.r.squared - summary(rt_recog_mean_f_seg_1)$adj.r.squared
```

Estimates from best fitting models

```{r rt fitted estimates, include=TRUE, warning=FALSE, message=FALSE}
# Slopes
# report coefficient values for s2_age in paper (estimate, SE, t, p)
summary(rt_cs_mean_m_lm)
summary(rt_cs_mean_f_lm)

summary(rt_recog_mean_m_lm)
summary(rt_recog_mean_f_lm)
```

# Software citations

```{r cite, include=TRUE, warning=FALSE, message=FALSE}

# R
citation()

# R version
version$version.string

# segmented package
citation("segmented")

# segmented package version
packageVersion("segmented")
```

Summary of results

On accuracy (cs and recog), best fitting model is linear for males 
and 1-changepoint for females

On rts (cs and recog), the best fitting models are linear for males and females